Learning Actionable Analytics (with applications for reducing defects and reducing runtimes)

### 2015-05-12 how.tex
Updates from ShareLaTeX

\documentclass[conference]{IEEEtran}
%related work (gentling them in)
%future work: local patch operator
%complexity
% planning in ai

@@

 This paper uses  HOW to learn methods
 to reduce defects and decrease program runtime.
 For the test data of this paper, those improvements   reduce
 defect counts and  runtimes by   
 (25,33)\% (median)   to (37,78)\% (max) of
 their original values.
  \end{abstract}
  \begin{IEEEkeywords}
Defect prediction, configuration, prediction, planning, case-based reasoning.
  \end{IEEEkeywords}

\section{Introduction}
Business  users   now demand   data mining tools
that  support a  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
and follow-up
that occurs {\em after} the mining, rather than just  the mining itself. For example:
\bi
\item 
Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users might now demand a {\em plan} to
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
a {\em plan} to reduce that runtime.
\ei
In this paper, we say   a ``plan'' is a  set of changes
proposed  in order to better achieve some goal. 
Specifically, for every $Z$  that requests a plan $\delta$ to
  adjust   attributes $Z_j$:
\begin{equation}\label{eq:one}
Z_j = Z_j + \delta_j
\end{equation}
The usual way such plans are generated in the SE literature
is via some search-based method. Recently, we have made  some progress
in this area with the  GALE optimizer~\cite{krall14}. 
Like other search-based algorithms,
 GALE executes by proposing multiple $\delta$s
 (one for each generation in an evolutionary program).
 In practice, GALE makes
 extensive changes to initial sample of project data
 (perhaps 
100s to 100,000s of mutations). Each of these mutations
is reassessed using some domain model.
GALE uses an eigenvector technique to find better mutants faster than standard search-based SE tools~\cite{deb00a,zit02}.

When presented to some  software project managers, the GALE approach was challenged by
two business users. Note that the   issues they raise are
challenges not just  about GALE, but also  about any     optimization algorithm  that  
 make extensive use of mutation  and domain models.
One  user asked:
\begin{quote}
ISSUE~\#1: {\em ``What if I only want the minimum useful change, not
the maximum possible?''}
\end{quote}
Here, the business user is asking for solutions  that do not require extensive 
changes to their current  approaches.

Later in that discussion, another business user asked:
\begin{quote}
ISSUE~\#2: {\em ``But what if I do not have a domain model of my business?''}
\end{quote}
This also is a legitimate issue.
As discussed below,
building (or accessing) and certifying and maintaining a 
trust-worthy model  can  takes  long time, especially in software engineering.

This paper addresses  ISSUE~\#1 and ISSUE~\#2 using case-based reasoning.
CBR  is purely based on the direct
adaptation of previous cases based on the similarity of
those cases with the current situation. That is, 
CBR based system have no explicit world model.
Rather, that model is implicit within  the available past
cases in the case cache.

This paper presents a novel CBR tool called ``HOW''. This tool  restricts test data $\delta$s to {\em minimal
displacements} (very small changes to a subset of the attributes)
thereby addressing ISSUE \#1. 
HOW ensures that the   $\delta$ displaces test data towards some region in the training data.
This is important since we address ISSUES~\#2 via some data mining
tool that train  a predictor from the training  data. Since we are pushing test data
into regions of the training data,  we can use the predictor to assess
the effectiveness of the $\delta$.
This paper shows that  that HOW's plans can be very effective indeed.
In the experiments shown below, HOW's plans
reduced the  expected
value of:
\bi
\item The  defects found  in Jureczko et al.'s    object-oriented JAVA systems~\cite{jureczko10}.
\item The runtime of the  software systems   configured by  Seigmund et al.~\cite{sven12}.
\ei 
In summary, this paper makes five contributions:
\be
\item The HOW  case-based planner (which has not been presented before).
\item The experiments of this paper showing that HOW planner is useful for real-world tasks;
e.g. reducing the
expected values of the Jureczko software defects and the Seigmund software runtimes.
\item Apart from HOW, we  also offers a general framework for using  predictive system to build planners;
\item Further, we offer a  test to recognizes  when we should {\em not}  plan: specifically,
planning is
not recommended in domains with  poorly performing predictors;
\item Last, we  offer a general verification rig for assessing the value of plans generated within this rig.
\ee

\section{Background and Related Work}
\subsection{Search-based Software Engineering (SBSE)} CPUs are becoming
faster and cloud computing enables access to more CPUs. Hence, it is now pratical
to solve large problems in SBSE via 
genetic algorithms  that build multiple generations by  mutating  an existing population,  
the combine the best mutants to fond the next population~\cite{Harman2009,Harman2011}. Examples of these algorithms include GALE, NSGA-II, NSGA-III, SPEA2, IBEA, particle swarm optimization, MOEA/D, etc.~\cite{krall14,deb00a,zit02,zit04,%
deb14,Cui2005a,zhang07:TEC}.

SBSE methods can make extensive mutations to the data they are exploring. Hence, these methods
require some domain model that can assess new and highly novel examples.

@@

 HOW    is very different
to the above CBR systems.
Firstly, HOW is a {\em planner} while the above
CBR methods are all {\em prediction} algorithms\footnote{The disctiniction being that
predictions systems tell you ``what is'', while planners tell you ``what to do''.}.
Secondly, all the above CBR tools explore the nearest ``k'' neighors,
HOW, on the other hand,  explores the {\em gradient between the scores of pairs of nearest neighbors}. 
To put that
another way,   the above CBR methods explore {\em points} while
HOW explores the slope {\em between} points.

\subsection{Prior Work}
HOW is the third
generation of our experiments with planning and CBR.
Generation one of that work resulted in
 W2, a CBR planner that reflected over the delta of raw  attributes~\cite{6600685}. It suffered from all-too-frequent {\em optimization failure}; i.e. when its plans were applied the resulting observed
performances {\em increased} in up to $\frac{1}{3}$rd of test cases.

@@

@@

\begin{figure}[t!]
\small
~\hrule~

@@

Those same regions can be used for case-based 
{\em planning}. However,  instead of leaving test cases in the regions
found during training, we {\em displace} them towards regions
with better objective scores.
To achieve this, 
during training, HOW clusters data using the procedure of \fig{where}a,
then draws {\tt Slope}s between the centroids of  clusters $X,Y$ that are nearest
neighbors.
By convention, we say that the cluster containing $X$ has the  better objective scores. This means we can say that the {\tt Slope} is a useful direction for displacing
data; i.e.  away from $Y$ and towards $X$ (for an implementation of such a {\tt Slope} class, see \fig{howcode}, line~1).
 During testing,  HOW   finds the  nearest {\tt Slope} to every test case. That  test case is   then displaced towards the better end
of that {\tt Slope}; i.e. towards  $X$ (see \fig{howcode}, line 51).

 HOW needs to  determine the  exact {\em direction} and {\em magnitude} of 
displacement.  To address  ISSUE~\#1,
HOW uses two techniques to result in 
{\em minimal displacements} of project data. Firstly,
HOW sorts  the independent variables
using the procedure of  \fig{where}d. It then only
displaces the    best $B$\%  of that sort, while leaving the rest untouched
(see \fig{howcode}, line 38).
In  the experiments shown below,  we explore   \mbox{$B \in \{25,50,100\}\%$}.

@@

Turning now to the displacement {\em direction},
our experiments show that it useful to  carefully reflect on exactly where HOW should  push each test case.
HOW replaces each cluster   with one {\em exemplar} that best
characterizes the data in that cluster. It then displaces the
test data towards that exemplar, as shown in \fig{howcode}, line 20. In summary, 
If  $Z_j$ is an independent variable within the
test case $Z$ and  $X$ is the exemplar found at training time,
then we displace $Z$ as follows:
\[Z_j =  Z_j + F*(X_j - Z_j)\]
Here, $F$ is the  magnitude controller, discussed above. Note that this
this equations satisfies our definition of a ``plan'' as offered in
\eq{one} in the  Introduction.

@@

However, for the Seigmund data, we found a better exemplar generation policy.
\fig{pp} shows the effects of selecting different kinds of exemplars for 
Berkeley DB, one of the Seigmund data sets.
The \textcolor{black}{\bf black baseline} curve  shows the runtimes for 600 random configurations
of the Makefiles used in Berkely DB. The other two curves shows the predictions of the runtimes if test cases are displaced towards exemplars
  built from:
\bi
\item The \textcolor{red}{\bf   mean}    value of all cases in the  cluster.
\item The  case with the \textcolor{green}{\bf  best}  objective score. 
\ei
Note that using the best as the exemplar does better than useing the mean.
The pattern shown in \fig{pp} has been seen in all the Seigmund data sets;
i.e. for data sets with continuous classes,
best results come from using exemplars formed from the best instance (and not the mean).

@@

Some organizations have the resources to 
run repeated trials to assess the effect of different treatments.
For example, in one remarkable recent study, Bente et al. report results
were the same specification was developed into a product by four different organizations~\cite{Anda2009}. Given those kind of resources, it would be possible
to (say) take a code base, assign it to $N$ different teams, apply
HOW's plans the code used by  $N/2$ teams, then check in 12 months time
if those halves of the teams have fewer defects than the other.

Very few industrial or research groups have access
to the kinds of resources using by Bente et al. (evidence: in the size years since the
publication of that work, we know of only one of two similar studies). Hence, it
we wish to verify HOW's plans, then some other framework is required that is less
resource intensive.

We propose using prediction systems (built by data mining) as a verification
tool for planners. In domains with data, from which data miners can build
predictors of adequate performance, then those predictors can assess the value
of the changes proposed by HOW.

Turning now to the {\em test} data, this is passed to the { predictor}
to generate a baseline {\em before} performance score.
Note that if our { predictors} fail to perform adequately on the test data,
then we cannot trust them to comment on our plans. Accordingly,
if that performance is unsatisfactory, we abort.
Else, we (1)~apply the { planner} to alter the {\em test} data;
then (2)~apply the { predictor} to the altered data {\em test'};
then (3)~return data on the {\em before,after} predictions.

@@

\subsection{When Not to Plan: the Problem of Poor Predictors}

@@

The ``tuned'' columns of \fig{j} show that we can salvage some of the data sets.
 Pelayo and Dick~\cite{pelayo07} report that defect prediction can be improved via SMOTE~\cite{Chawla2002}; i.e. an over-sampling of minority-class examples.
 Also, Fu et al.~\cite{fu:ase15} report that parameter tuning with differential evolution~\cite{storn97}
applied to defect predictors can quickly explore the tuning
options of Random Forest to find better settings for the size of the forest, the termination criteria
for tree generation, etc.
The rows marked with $\star$ in \fig{j} show data sets whose performance was improved remarkably by these
techniques. For example, in {\em poi}, the recall increased by 10\% while the false alarm rate dropped by 20\%.
However,  as might be expected
the data sets in the  three bottom rows remain
unsalvageable.

@@

\subsection{Data}
To assess HOW, we applied its plans to all the Seigmund data, and the five ``good''
  Jureczko   data sets {\em jedit, ivy, ant, lucene} and {\em poi}.
  The training and tests used for Jurecko were display in \fig{j}. For these data
  sets, training was managed by a Random Forest (tuned in the SMOTED training data using differential
  evolution). \fig{ck} shows the attributes in the Jureczko data.

  The Seigmund data sets are far more varied than the Jurecko data.
  These contain the results of applying a large number of mutators
  to the configuration options of the Makefiles of  BerkeleyDB (the Java  and  Windows  versions),  Apache,  SQLite,  LLVM, and x264. For details on those datasets, see \fig{cpm}.

@@

@@

\begin{figure}[!t] 
  {\scriptsize \textbf{Ant}\\[0.1cm]}
  {\scriptsize  \begin{tabular}{{l@{~~~~}l@{}r@{~~~~}r@{~~~~~}c@{}r}}
      \arrayrulecolor{darkgray}
      \rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} && \textbf{Bugs}\\
  1 &  0.75-weight &    0.71  &  0.19 & \quart{0}{25}{10}{49} & 62\\
  1 &         0.75 &    0.72  &  0.14 & \quart{6}{19}{12}{49} & 64\\
\hline  2 &          0.5 &    0.86  &  0.08 & \quart{25}{11}{31}{49} & 77 \\
  2 &   0.5-weight &    0.88  &  0.14 & \quart{20}{19}{33}{49} & 77\\
\hline  3 & 0.5-weight-B=50\% &    0.91  &  0.08 & \quart{32}{11}{37}{49} & 80\\
  3 & 0.75-weight-B=50\% &    0.92  &  0.09 & \quart{31}{12}{39}{49} & 82\\
  3 &  0.25-weight &    0.92  &  0.04 & \quart{36}{5}{39}{49} & 81\\
\hline  4 &         0.25 &    0.93  &  0.05 & \quart{37}{7}{40}{49} & 83\\
  4 & 0.75-weight-B=25\% &    0.94  &  0.09 & \quart{36}{12}{41}{49} & 82\\
\hline  5 & 0.25-weight-B=50\% &    0.97  &  0.03 & \quart{44}{4}{45}{49} & 86\\
  5 & 0.5-weight-B=25\% &    0.98  &  0.06 & \quart{40}{8}{47}{49} & 86 \\
\hline  6 & 0.25-weight-B=25\% &    0.99  &  0.01 & \quart{48}{1}{48}{49} & 87\\
\hline  7 &         Base &    1.0  &  0.0 & \quart{49}{0}{49}{49} & 90\\
\hline \end{tabular}}
%\end{minipage}\\[0.1cm]

%\noindent\begin{minipage}{0.5\textwidth}
 % \flushleft
 \vspace{2mm}
 \noindent {\scriptsize \textbf{Ivy}\\[0.1cm]}
{\scriptsize  \begin{tabular}{{l@{~~~~}l@{}r@{~~~~}r@{~~~~}c@{}r}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & & \textbf{Bugs}\\
  1 &  0.75-weight &    0.75  &  0.18 & \quart{0}{29}{8}{49} & 50\\
  1 &         0.75 &    0.79  &  0.12 & \quart{8}{20}{14}{49} & 53\\
\hline  2 &          0.5 &    0.84  &  0.1 & \quart{18}{16}{23}{49} & 56\\
  2 &   0.5-weight &    0.84  &  0.08 & \quart{14}{14}{23}{49} & 56\\
  2 & 0.75-weight-B=50\% &    0.84  &  0.09 & \quart{18}{15}{23}{49} & 56 \\
\hline  3 & 0.5-weight-B=50\% &    0.91  &  0.05 & \quart{29}{9}{34}{49} & 61\\
\hline  4 & 0.25-weight-B=50\% &    0.96  &  0.01 & \quart{43}{1}{43}{49} & 64\\
  4 & 0.75-weight-B=25\% &    0.96  &  0.05 & \quart{39}{9}{43}{49} & 65\\
  4 &         0.25 &    0.97  &  0.01 & \quart{43}{1}{44}{49} & 65\\
  4 &  0.25-weight &    0.97  &  0.05 & \quart{39}{9}{44}{49} & 65 \\
  4 & 0.5-weight-B=25\% &    0.97  &  0.03 & \quart{43}{5}{44}{49} & 65 \\
\hline  5 & 0.25-weight-B=25\% &    0.99  &  0.01 & \quart{48}{1}{48}{49} & 66\\
\hline  
6 &         Base &    1.0  &  0.0 & \quart{49}{0}{49}{49} & 67\\
\hline \end{tabular}}
%\end{minipage}\\[0.1cm]

>>>>

\documentclass[conference]{IEEEtran}
%future work: local patch operator
%complexity

@@

 This paper uses  HOW to learn methods
 to reduce defects and decrease program runtime.
 For the test data of this paper, those improvements   reduce
 the expected values of defect counts and  runtimes to    
 72\%,75\%  (median) and 66\%,20\% (best case) of the initial values.
  \end{abstract}
  \begin{IEEEkeywords}
Defect prediction, configuration, prediction, planning, case-based reasoning.
  \end{IEEEkeywords}

\section{Introduction}
Suppose someone suggests a change to your project. How might you assess the effectiveness of that change?

This is an open and pressing issue.
Business  users   now demand   data mining tools
that  support a  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
and follow-up
that occurs {\em after} the mining, rather than just  the mining itself; e.g.
\bi
\item 
Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users might now demand a {\em plan} to
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
a {\em plan} to reduce that runtime.
\ei
How to assess such {\em plans}?
Some organizations have the resources to 
run repeated trials to assess  different treatments.
For example, in one remarkable recent study, Bente et al. report results
were the same specification was developed into a product by four different organizations~\cite{Anda2009}. Given those kind of resources, it would be possible
to (say) take a code base, assign it to $N$ different teams, apply
HOW's plans the code used by  $N/2$ teams, then check in 12 months time
if those halves of the teams have fewer defects than the other.

Very few industrial or research groups have access
to the kinds of resources using by Bente et al. (evidence: in the size years since the
publication of that work, we know of only one of two similar studies). Also, given the
diversity of modern software projects, it might be unreasonable to demand that all
proposed changes for all projects are always evaluated by something like the Bente et al. study.

This paper proposes another, less resource intensive,  method for assessing the effectiveness of proposed
changes to a project.  
We propose using prediction systems (built by data mining) as a verification
tool for planners. In domains where  data miners can build
predictors of adequate performance, then those predictors can assess the value
of proposed changes.

The planner explored in this paper is the ``HOW'' case-based planner.
HOW's  ``plans'' are  set of changes
proposed  in order to better achieve some goal. 
Specifically, for every test case $Z$, HOW proposes a  plan $\delta$ to
  adjust   attributes $Z_j$:
\begin{equation}\label{eq:one}
Z_j = Z_j + \delta_j
\end{equation}
For example, to simplify a  large bug-prone  method, HOW might suggest
to a developer to reduce its size (i.e.  refactor that code by, say, splitting it across
two simpler functions).

HOW's  plans displace the test case into some region
of the training data. This is important since we will use that training data
to build a predictor to assess the impact of that plan.
To put that another way, when a novel test case is presented,
HOW says 
\begin{quote}{\em ``I may not know too much about that, but once you change into something like
what I've already seen, then I can suggest how to change it''.}
\end{quote}
This paper shows experimentally  that   HOW's plans can be quite effective and 
 significantly reduce the  predicted
value of:
\bi
\item The  defects found  in Jureczko et al.'s    object-oriented JAVA systems~\cite{jureczko10}.
\item The runtime of the  software systems   configured by  Seigmund et al.~\cite{sven12}.
\ei 
The rest of this paper presents the background to this work in Section~2; the design of HOW in Section~3;
then our experimental results  in Section~4. Those experiments show that HOW can   reduce the expected
value of defects in Jureczko   data sets by 72\% to 66\%  (median,best).
Also, in the Seigmund data sets, HOW can   reduce the expected value of the runtimes by  75\% to 20\% (median,best).

\section{Background}
\subsection{Search-based Software Engineering (SBSE)}
One way to propose changes to software artifacts
is   via some search-based method~\cite{Harman2009,Harman2011}. Such SBSE methods are   evolutionary programs that 
make
 extensive changes to  some initial sample of project data
 (perhaps 
100s to 100,000s of mutations). Each of these mutations
is reassessed using some domain model.
Examples of these algorithms include GALE, NSGA-II, NSGA-III, SPEA2, IBEA, particle swarm optimization, MOEA/D, etc.~\cite{krall14,deb00a,zit02,zit04,%
deb14,Cui2005a,zhang07:TEC}.

These SBSE methods can make extensive mutations to the data they are exploring. Hence, these methods
require some domain model that can assess new and highly novel examples.

@@

\subsection{Prior Work}

HOW is the third
generation of our exploration of  planning and CBR and software engineering.
Generation one  resulted in
 W2, a CBR planner that reflected over the delta of raw  attributes~\cite{6600685}. It suffered from all-too-frequent {\em optimization failure}; i.e. when its plans were applied the resulting observed
performances {\em increased} in up to $\frac{1}{3}$rd of test cases.

@@

HOW is very different to other CBR systems. Firstly, HOW is a {\em planner} while many
CBR methods in software engineering explore  {\em prediction} algorithms\footnote{The distinction being that
predictions systems tell you ``what is'', while planners tell you ``what to do''.}.
Secondly, all the above CBR tools explore the nearest ``k'' neighors,
HOW, on the other hand, explores the {\em gradient between the scores of pairs of nearest neighbors}. 
To put that another way, the above CBR methods explore {\em points} while
HOW explores the slope {\em between} points.

\fig{howcode}, line~1, shows  an implementation of such a {\tt Slope} class and
the rest
of this section  details   {\tt Slope}'s role within the HOW system.

@@

\begin{figure}[b!]
\small
~\hrule~

@@

Those same regions can be used for case-based 
{\em planning}. However,  instead of leaving test cases in the regions
found during training, we {\em displace} them towards regions
with better objective scores.
To achieve this, 
during training, HOW clusters data using the procedure of \fig{where}a,
then draws {\tt Slope}s between the centroids of  clusters $X,Y$ that are nearest
neighbors.
By convention, we say that the cluster containing $X$ has the  better objective scores. This means we can say that the {\tt Slope} is a useful direction for displacing
data; i.e.  away from $Y$ and towards $X$. 
 During testing,  HOW   finds the  nearest {\tt Slope} to every test case. That  test case is   then displaced towards the better end
of that {\tt Slope}; i.e. towards  $X$ (see \fig{howcode}, line 51).

 HOW needs to  determine the  exact {\em direction} and {\em magnitude} of 
displacement.   
HOW uses two techniques to result in 
{\em minimal displacements} of project data. Those techniques are described below. Note
 that the net result of these techniques is to drive the test case into some region of the training 
 set (this is important since we will use a predictor learned from that training data to assess
 the effectiveness of HOW's plans).

 Firstly,
HOW sorts  the independent variables
using the procedure of  \fig{where}d. It then only
displaces the    best $B$\%  of that sort, while leaving the rest untouched
(see \fig{howcode}, line 38).
In  the experiments shown below,  we explore   \mbox{$B \in \{25,50,100\}\%$}.

@@

Turning now to the displacement {\em direction},
our experiments show that it useful to  carefully reflect on exactly where HOW should  push each test case.
HOW replaces each cluster   with one {\em exemplar} that best
characterizes the data in that cluster. It then displaces the
test data towards that exemplar, as shown in \fig{howcode}, line 20. In summary, 
If  $Z_j$ is an independent variable within the
test case $Z$ and  $X$ is the exemplar found at training time,
then we displace $Z_j$ as follows:
\[Z_j =  Z_j + F*(X_j - Z_j)\]
Here, $F$ is the  magnitude controller, discussed above.

@@

However, for the Seigmund data, we found a better exemplar generation policy.
\fig{pp} shows the effects of selecting different kinds of exemplars for 
Berkeley DB, one of the Seigmund data sets.
The \textcolor{black}{\bf black baseline} curve  shows the runtimes for 600 random configurations
of the Makefiles used in Berkely DB. The other two curves shows the predictions of the runtimes if test cases are displaced towards exemplars
  built from:
\bi
\item The \textcolor{red}{\bf   mean}    value of all cases in the  cluster.
\item The  case with the \textcolor{green}{\bf  best}  objective score. 
\ei
Note that using the best as the exemplar does better than useing the mean.
The general pattern shown of \fig{pp} has been seen in all the Seigmund data sets;
i.e. for data sets with continuous classes,
best results come from using exemplars formed from the best instance (and not the mean).

@@

@@

Turning now to the {\em test} data, this is passed to the { predictor}
to generate a baseline {\em before}  score.
If our { predictors} fail to perform adequately on the test data,
then we cannot trust them to comment on our plans. Accordingly,
if that performance is unsatisfactory, we abort.
Else, we (1)~apply the { planner} to alter the {\em test} data;
then (2)~apply the { predictor} to the altered data {\em test'};
then (3)~return data on the {\em before,after} predictions.

\subsection{When Not to Plan: the Problem of Poor Predictors}\label{sec:notplan}

@@

The ``tuned'' columns of \fig{j} show that we can salvage some of the data sets.
 Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e. an over-sampling of minority-class examples.
 Also, Fu et al.~\cite{fu:ase15} report that parameter tuning with differential evolution~\cite{storn97}
can quickly explore the tuning
options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
for tree generation, etc.
The rows marked with $\star$ in \fig{j} show data sets whose performance was improved remarkably by these
techniques. For example, in {\em poi}, the recall increased by 10\% while the false alarm rate dropped by 20\%.
However,  as might have been  expected
we could not salvage the data sets in the  three bottom rows.

@@

\subsection{Data}
We applied HOW's plans to all  Seigmund data, and the five ``good''
  Jureczko   data sets {\em jedit, ivy, ant, lucene} and {\em poi}.
  The training and tests used for Jurecko were display in \fig{j}. For these data
  sets, training was managed by a Random Forest (tuned in the SMOTED training data using differential
  evolution). \fig{ck} shows Jureczko data attributes.

   \fig{cpm} shows the  Seigmund data attributes.
   There were more diverse than the 
    Jurecko data since they
   contain the results of applying many  mutators
  to the configuration options of the Makefiles of  BerkeleyDB (the Java  and  Windows  versions),  Apache,  SQLite,  LLVM, and x264.

@@

   \subsection{Procedures}
  \fig{work} showed how to   build and test HOW's plans.
The    Seigmund and Jureczko data sets have continuous and discrete classes (respectively).
Hence, when building a predictor, we used (1)~Random Forest Regressors for the Seigmund data and (2)~Random Forests
for the Jureczko data.

   For both the Siegmind and Jureczko, we designed a procedure such we had at least
   20 samples from the \fig{work} procedure for  each data set (the value of 20 was choosen as the minimum use sample size via the central limit theorem):
   \bi
   \item
  For the Siegmind data, 
  25 times, 
  we randomized the order of the data, trained on one half (via
  Random Forest Regressor), then tested on the other.
  \item
  For the Jureczko, if we just  used the training and test sets of  \fig{j}, that would only
  give us sample per data set. Such a singleton sample could hide important details since the methods
  used for parameter tuning (SMOTE and differential evolution) use some stochastic sub-routines.
  Hence, for each  of the training sets of \fig{j}, we ran SMOTE+differential evolution 25 times to learn
  parameters for Random Forests.
  \ei
\fig{where} defined two control parameters for HOW:
\bi
\item  $F$   controls the size of the displacement;
\item   $B$   controls how many  attributes are displaced.
\ei
The   25 repeats defined above were executed for various values of $F,B$.
A Scott-Knott procedure was used to test for statistically significant differences in the results.
The Scott-Knott
procedure recommended by Mittas \& Angelis in their 2013
IEEE TSE paper~\cite{mittas13}.  This method
sorts a list of $l$ treatments with $ls$ measurements by their median
score. It then
splits $l$ into sub-lists $m,n$ in order to maximize the expected value of
 differences  in the observed performances
before and after divisions. E.g. for lists $l,m,n$ of size $ls,ms,ns$ where $l=m\cup n$:
 $E(\Delta)=\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2$.
Scott-Knott  applies an  hypothesis test $H$ to check
if $m,n$ are significantly different. If so, Scott-Knott  recurses on the splits.

@@

\begin{figure*}[!t]
\begin{center}
\begin{minipage}{.44\linewidth}
%\noindent\begin{minipage}{0.5\textwidth}
  {\scriptsize \textbf{Apache}\\[0.1cm]}
  {\scriptsize  \begin{tabular}{{llrrc}}
      \arrayrulecolor{darkgray}
      \rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 & F=0.75, B=25 &    0.79  &  0.06 & \quart{0}{11}{7}{49} \\
\hline  2 & F=0.75, B=50 &    0.82  &  0.07 & \quart{7}{14}{13}{49} \\
\hline  3 &  F=0.75, B=100  &    0.85  &  0.06 & \quart{11}{12}{19}{49} \\
  3 & F=0.5, B=25 &    0.87  &  0.03 & \quart{19}{6}{23}{49} \\
  3 & F=0.5, B=50 &    0.87  &  0.07 & \quart{17}{14}{23}{49} \\
\hline  4 &   F=0.5, B=100  &    0.91  &  0.07 & \quart{21}{14}{31}{49} \\
\hline  5 & F=0.25, B=25 &    0.92  &  0.05 & \quart{29}{10}{33}{49} \\
  5 & F=0.25, B=50 &    0.94  &  0.04 & \quart{35}{8}{37}{49} \\
  5 &  F=0.25,B=100   &    0.95  &  0.04 & \quart{35}{8}{39}{49} \\
\hline  6 &     Baseline &    1.0  &  0.0 & \quart{49}{0}{49}{49} \\
\hline \end{tabular}}

\vspace{2mm}\noindent {\scriptsize \textbf{BDBC}\\[0.1cm]}
{\scriptsize  \begin{tabular}{{llrrc}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 & F=0.75, B=50 &    0.2  &  0.03 & \quart{0}{1}{1}{49} \\
  1 & F=0.75, B=25 &    0.2  &  0.05 & \quart{0}{3}{1}{49} \\
\hline  2 &  F=0.75, B=100  &    0.22  &  0.07 & \quart{1}{5}{2}{49} \\
\hline  3 & F=0.5, B=25 &    0.39  &  0.06 & \quart{11}{4}{12}{49} \\
  3 & F=0.5, B=50 &    0.4  &  0.05 & \quart{12}{3}{13}{49} \\
\hline  4 &   F=0.5,B=100   &    0.45  &  0.08 & \quart{14}{4}{16}{49} \\
\hline  5 & F=0.25, B=25 &    0.65  &  0.05 & \quart{27}{3}{28}{49} \\
  5 & F=0.25, B=50 &    0.66  &  0.05 & \quart{27}{3}{29}{49} \\
\hline  6 &  F=0.25,B=100   &    0.69  &  0.05 & \quart{29}{3}{31}{49} \\
\hline  7 &     Baseline &    1.0  &  0.0 & \quart{49}{0}{49}{49} \\
\hline \end{tabular}}
%\end{minipage}

\vspace{2mm}\noindent{\scriptsize \textbf{BDBJ}\\[0.1cm]}
{\scriptsize  \begin{tabular}{{llrrc}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 & F=0.75, B=25 &    0.64  &  0.1 & \quart{1}{12}{6}{49} \\
  1 & F=0.75, B=50 &    0.68  &  0.14 & \quart{0}{17}{10}{49} \\
\hline  2 &  F=0.75, B=100   &    0.74  &  0.1 & \quart{12}{12}{18}{49} \\
  2 & F=0.5, B=25 &    0.76  &  0.09 & \quart{15}{11}{20}{49} \\
  2 & F=0.5, B=50 &    0.78  &  0.17 & \quart{15}{21}{23}{49} \\
  2 &   F=0.5,B=100   &    0.8  &  0.07 & \quart{20}{9}{25}{49} \\
\hline  3 & F=0.25, B=25 &    0.85  &  0.05 & \quart{30}{6}{31}{49} \\
\hline  4 & F=0.25, B=50 &    0.89  &  0.09 & \quart{30}{11}{36}{49} \\
  4 &  F=0.25, B=100  &    0.9  &  0.06 & \quart{35}{7}{37}{49} \\
\hline  5 &     Baseline &    1.0  &  0.0 & \quart{49}{0}{49}{49} \\
\hline \end{tabular}}
%\end{minipage}
%\noindent\begin{minipage}{0.5\textwidth}
%  \flushleft
\end{minipage}~~~~~\begin{minipage}{.44\linewidth}
\noindent
{\scriptsize \textbf{LLVM}\\[0.1cm]}
  {\scriptsize  \begin{tabular}{{llrrc}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 & F=0.75, B=25 &    0.92  &  0.02 & \quart{0}{11}{5}{49} \\
  1 & F=0.75, B=50 &    0.93  &  0.02 & \quart{5}{11}{11}{49} \\
  1 &  F=0.75, B=100   &    0.93  &  0.02 & \quart{11}{11}{11}{49} \\
\hline  2 & F=0.5, B=25 &    0.95  &  0.01 & \quart{16}{6}{22}{49} \\
  2 & F=0.5, B=50 &    0.95  &  0.01 & \quart{16}{6}{22}{49} \\
  2 &   F=0.5, B=100   &    0.96  &  0.01 & \quart{22}{5}{27}{49} \\
\hline  3 & F=0.25, B=25 &    0.97  &  0.01 & \quart{33}{5}{33}{49} \\
  3 & F=0.25, B=50 &    0.98  &  0.0 & \quart{38}{0}{38}{49} \\
  3 &  F=0.25, B=100  &    0.98  &  0.01 & \quart{38}{6}{38}{49} \\
\hline  4 &     Baseline &    1.0  &  0.0 & \quart{49}{0}{49}{49} \\
\hline \end{tabular}}
%\end{minipage}

\vspace{2mm}\noindent {\scriptsize \textbf{SQL}\\[0.1cm]}
  {\scriptsize  \begin{tabular}{{llrrc}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 & F=0.75, B=25 &    0.98  &  0.0 & \quart{0}{0}{0}{49} \\
  1 & F=0.75, B=50 &    0.98  &  0.0 & \quart{0}{0}{0}{49} \\
  1 & F=0.75, B=100 &   0.99  &  0.01 & \quart{0}{24}{24}{49} \\
  1 & F=0.5, B=25 &    0.99  &  0.0 & \quart{24}{0}{24}{49} \\
  1 & F=0.5, B=50 &    0.99  &  0.0 & \quart{24}{0}{24}{49} \\
  1 & F=0.5, B=100 &    0.99  &  0.01 & \quart{24}{25}{24}{49} \\
  1 & F=0.25, B=50 &    1.0  &  0.0 & \quart{49}{0}{49}{49} \\
  1 & F=0.25-weight-B=25 &    1.0  &  0.0 & \quart{49}{0}{49}{49} \\
  1 & F=0.25, B=100 &    1.0  &  0.0 & \quart{49}{0}{49}{49} \\
  1 &     Baseline &    1.0  &  0.0 & \quart{49}{0}{49}{49} \\
  \hline \end{tabular}}
%\end{minipage}
%\noindent\begin{minipage}{0.5\textwidth}
%  \flushleft
